---
title: "FINAL STAT 1601 project"
author: "Isha Anand and Aditi Mahabal"
date: "4/30/2020"
output: pdf_document
---
```{r}
library(ggplot2)
library(dplyr)
library(ISLR)
library(MASS)
library(RColorBrewer)
```


Data Wrangling
```{r}
#This formula for removing all unknown values and fixing factor data was taken and adapted to our uses from the site cited in the references

police2 <-read.csv("/Users/ishaanand/Downloads/STAT 1601/Data Sets/POLICEKILL - PoliceKillings.csv",
                   stringsAsFactors = FALSE,na.strings = c("NA", "N/A", "Unknown", "NULL", ".P"))

pk<- droplevels(na.omit(police2))

#Eliminated all factors from variables that were hindering our analysis
```

Research Question 1: Gender Two-Sample Inference of Proportions
```{r}
pkGender<-pk%>%
  dplyr::select(gender)

table(pkGender)
#female = 20/410, male = 390/410

prop.test(c(390, 20), c(410, 410), conf.level=0.95)
prop.test(c(390, 20), c(410, 410), alternative="greater")
```

Research Question 2: Gender and Age Two-Sample Inference of Means
```{r}


genderAge<-pk%>%
  dplyr::select(age, gender)

ggplot(genderAge, mapping=aes(x=gender, y=age,))+
  geom_boxplot(color="black", fill="pink",  outlier.color = "blue", outlier.alpha=.5)+
  labs(x="Gender", y="Age", title= "Victim Gender vs. Age")

t.test(age~gender, genderAge, conf.level=0.95)

t.test(age~gender, genderAge, alternative="less")
```

Research Question 3: Single Inference: Mean household income of those killed by the police
```{r}
h_income1<-as.numeric(pk$h_income)
t.test(h_income1, conf.level = 0.95 ) 
```


Research Question 4: Multiple Inference: Proportions of Race/Ethnicites Killed
```{r}
chisq.test(table(pk$raceethnicity))
```
#Bar Graph showing the difference in proportions of the race/ethinicites group
```{r}
ggplot(pk,mapping=aes(x=raceethnicity))+geom_bar(fill=brewer.pal(5,"Set2"))+
  theme(axis.text.x=element_text(angle=90))+
  labs(x="Race/Ethnicity", y="Number of Killings", title = "Number of Each Race/Ethnicity Group")
```

#Table of individuals under the age of 30 being killed, grouped by race/ethnicity
```{r}
pk4<-pk%>%
group_by(raceethnicity)%>%
  filter(age<30)
table(pk4$raceethnicity)
```

Research Question 5: Multiple Regression
```{r}
pk[,32]<-as.numeric(pk$pov)

pk%>%
  dplyr::select(urate, age, college, pov)%>%
  cor()


regression<-lm(age~urate+college+pov, pk)
steppedReg<-stepAIC(regression, direction="both", trace=FALSE)
summary(steppedReg)


ggplot(pk, mapping=aes(x=urate, y=age))+geom_point()+geom_smooth(method="lm", se=FALSE)+
  labs(x="Unemployment Rate", y="Age", title= "Victims' Community Unemployment Rate vs. Age")

ggplot(pk, mapping=aes(x=college, y=age))+geom_point()+geom_smooth(method="lm", se=FALSE)+
  labs(x="Amount of College Degree Holders", y="Age", title= "Victims' Community College Degree Holder % vs. Age")

ggplot(pk, mapping=aes(x=pov, y=age))+geom_point()+geom_smooth(method="lm", se=FALSE)+
  labs(x="Poverty Rate", y="Age", title= "Victims' Community Poverty Rate vs. Age")

```



#Data Exploration and Vizualization: Additional graphs and tables of variables

TABLES

Table 1: The number of each race/ethnicity group to be killed while unarmed 
```{r}
pk3<-pk%>%
group_by(raceethnicity)%>%
  filter(armed=="No")
table(pk3$raceethnicity)
```

Table 2: Average age of individuals killed in each state
```{r}
aggregate(pk[,2],list(pk$state),mean)
```

GRAPHS

Graph 1:Boxplot
```{r}
ggplot(pk, aes(x=raceethnicity, y=age,fill=raceethnicity))+geom_boxplot()+
  labs(x="Race/Ethnicity", y="Age", title = "Distribution of Race/Ethnicity vs. Age")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
```

Graph 2: Bubble Scatter Plot
```{r}
ggplot(data = pk, mapping = aes(x = state, y = college, size=age, fill=gender)) +
geom_point(alpha=0.5, shape=21, color="black") + scale_size(range = c(.1, 6))+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))+
  labs(x="State", y="College", title = "College Completion Rates per State")
```

Graph 3: Histogram
```{r}
ggplot(data = pk, mapping = aes(x=age, fill=raceethnicity)) + geom_histogram(bins=21,color="black")+
  labs(x="Age", y="Number of Killings", title = "Distribution of Ages vs. Number of Killings")
```

Graph 4&5: Bar Graph
```{r}
ggplot(pk, mapping=aes(x=gender, fill=gender))+geom_bar()+
  labs(x="Gender", y="Number of Victims", title= "Number of Male vs. Female Victims of Police Killings")

ggplot(pk, mapping=aes(x=state, fill=raceethnicity))+geom_bar(width=.8, alpha=0.8)+
  theme(axis.text.x=element_text(angle=90))+labs(x="State", y="Number of Killings", title= "Number of Killings per U.S. State")
```

Graphs 6,7,8: Density Plot
```{r}
ggplot(pk, mapping=aes(x=college, fill=raceethnicity))+geom_density(alpha=0.3, outline.type="full")+ 
  labs(x="Amount of College Degree Holders", y="Density", title= "Victim Distribution over Community % of College Degree Holders")

ggplot(pk, mapping=aes(x=urate, fill=raceethnicity))+geom_density(alpha=0.3, outline.type="full")+
  labs(x="Unemployment Rate", y="Density", title= "Distribution of Victims over Varying Community Unemployment Rates")

ggplot(pk, mapping=aes(x=pov, fill=raceethnicity))+geom_density(alpha=0.3, outline.type="full")+
  labs(x="Poverty Rate", y="Density", title= "Distribution of Victims over Varying Community Poverty Rates")
```

